---
title: "new tmap modes"
output: 
  bookdown::html_vignette2:
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  comment = "#>"
)
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(more, x[lines], more)
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

```



```{r, echo = FALSE, message = FALSE}
library(tmap)
library(tmap.mapgl)
```

Recall that *tmap* offers two modes: `"plot"` for static mapping and `"view"` for interactive mapping. See [introduction](https://r-tmap.github.io/tmap/articles/basics_modes). The `"view"` mode uses the JavaScript library Leaflet as backend.

The extension package *tmap.mapgl* offers two new modes which are also interactive: `"mapbox"` and `"maplibre"` which use the JavaScript libraries Mapbox GL and Maplibre GL respectively. An API key is requires to use `"mapbox"` (free for personal use), but `"maplibre"` is (as the name suggestions) free for any use.

Note that *tmap.mapgl* is a bridge between the R packages *mapgl* and *tmap*. It makes the functionality of *mapgl* (making the JavaScript libraries available to R) also available via the *tmap* user interface. 



## Switching modes

Get the current mode

```{r}
tmap_mode()
```

For this example we'll create a choropleth of well being per country. We assign the map to `map` without showing it. 

```{r}
map = tm_shape(World) + 
  tm_polygons("well_being",
    fill.scale = tm_scale_continuous(values = "pu_gn")) +
  tm_crs("auto")
```

Now we'll show the map in all (now four) modes

### The modes

```{r, fig.height = 3.5}
tmap_mode("plot")
map
```



```{r, fig.height = 3.5}
tmap_mode("view")
map
```


```{r, fig.height = 3.5}
tmap_mode("mapbox")
map
```

```{r, fig.height = 3.5}
tmap_mode("maplibre")
map
```




