---
title: "3D polygon layer"
output: 
  bookdown::html_vignette2:
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  comment = "#>"
)
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(more, x[lines], more)
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

```

```{r, echo = FALSE, message = FALSE}
library(tmap)
library(tmap.mapgl)
```


**tmap.mapgl** also features a new layer type, `tm_polygons_3d`, which is only available for the `"mapbox"` and `"maplibre"` modes.

This map layer is the same as `tm_polygons`, with one addition: polygons can be extruded into 3D shapes. The visual variable to control this is called `height`.



## Example: buildings heights

```{r}
tmap_mode("maplibre")

# get vector buildings
library(osmdata)
buildings <- opq(bbox = "Marina Bay, Singapore") |>
	add_osm_feature(key = "building") |>
	osmdata_sf()

# prepare levels column, assuming 2 if NA
library(dplyr)
buildings_poly <- buildings$osm_polygons |>
	mutate(levels = as.numeric(`building:levels`),
		   .before = 1) |>
	mutate(levels = if_else(is.na(levels), 2, levels),
		   height = levels * 3.5) # rough height estimate in meters


```

For the time being, we need to compute the maximum building height, because `tm_scale_asis` doesn't support units yet.

```{r, fig.height = 6}
# maximum building height
mx = max(buildings_poly$height)

# plot
tm_shape(buildings_poly) +
	tm_polygons_3d(height = "height", options = opt_tm_polygons_3d(height.max = mx)) + 
	tm_maplibre(pitch = 60)
```

## Example: population density data 

```{r, fig.height = 6}
NLD_dist$pop_dens = NLD_dist$population / NLD_dist$area

tm_shape(NLD_dist) +
  tm_polygons_3d(height = "pop_dens",
    fill = "edu_appl_sci",
    fill.scale = tm_scale_intervals(style = "kmeans", values = "-pu_gn"),
    fill.legend = tm_legend("University degree")) +
tm_maplibre(pitch = 45)
```

